---
title: "Find pooling region"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Find pooling region}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(findpoolreg)
library(dplyr)
```

## Preliminaries and Model Assumptions

The tests are based on the assumption that the annual/seasonal precipitation extremes $M_d^{(t)}$, observed at 
location $d$ in year $t$, follow a generalised extreme value (GEV) distribution with parameters
$$
	\mu_d(t) = \mu_d \exp\left(\frac{\alpha_d \mathrm{GMST}(t)}{ \mu_d} \right), 
	\quad
	\sigma_d(t) = \sigma_d \exp\left(\frac{\alpha_d \mathrm{GMST}(t)}{ \mu_d} \right), 
	\quad
	\gamma_d(t) =\gamma_d.
$$
The parameter vector is thus $\mathbf \vartheta_d = (\mu_d, \sigma_d, \gamma_d, \alpha_d)^\top$.
The null-hypothesis that is tested is for a given set of locations $A$ with $|A| \geq 2$ is
$$ H_0(A): \quad \exists\, \mathbf \vartheta_A \in (0, \infty)^2 \times \mathbb{R}^2:\  \forall\, d \in A: \quad \mathbf \vartheta_d = \mathbf \vartheta_A. $$
Therefore, whenever the test rejects the null-hypothesis for a pair, the pair can be considered as not homogeneous and observations from 
the respective locations should not be pooled to a larger sample. 

## Application of bootstrap procedure

We are going to apply the bootstrap procedures on the example data set contained in this package. 
Note that the example data set is a simulated data set, with  parameters
```{r, echo = FALSE}
data.frame(Location = 1:9, mu = c(23, 23, 22, 20, 20, 20, 20, 20, 21), 
           sigma = c(6, 6, 6, 5, 5, 6, 5, 5, 6), 
           gamma = c( 0.12, 0.1, 0.15, 0.1, 0.1, 0.15, 0.11, 0.1, 0.05), 
           alpha = c(rep(3, 3), rep(2.5, 6)))
```
We therefore have $\mathbf{\vartheta}_4 = \mathbf{\vartheta}_5 = \mathbf{\vartheta}_8$, 
and $\mathbf{\vartheta}_7$ deviates only very slightly.

First, we load the data and grid coordinates.
```{r}
data("example_grid")
data("example_data")
head(example_grid)
head(example_data)
```
We can have a look at the grid to get an impression of the region.
```{r}
leaflet::leaflet(data = example_grid, width = 500, height = 400)  %>% 
  leaflet::addTiles() %>%
  leaflet::addRectangles(lng1 = ~from_lon, lng2 = ~to_lon, lat1 = ~from_lat, lat2 = ~to_lat) %>% 
  leaflet::addLabelOnlyMarkers(~meanlon, ~meanlat, label =  ~as.character(Region), 
                               labelOptions = leaflet::labelOptions(noHide = TRUE, textOnly = TRUE,
                               textsize = 5, style = list("color" = "blue",  "font-weight" = "bold", 
                                                          "font-size" = "30px")))
```

Now, let's assume we want to find all regions for which the parameters of the scale-GEV distribution seem
sufficiently similar to those of the region with the label 5. 
To do so, we first apply the bootstrap procedure based on max-stable processes to all pairs $(5, d), d = 1, \ldots, 4, 6, \ldots, 9$.
We first make a list containing all those pairs and get values of the temporal covariate (here: 4-year smoothed GMST anomaly). 
For this, we assume that our data are observations from Years 1947 to 2021. 
Also, we check whether the column names of the data matrix are those used when defining the pairs (i.e. numbered from 1 to 9). 
```{r}
colnames(example_data)
subsets <- purrr::map(c(1:4, 6:9), ~ c(5, .x))
subsets[1:2]  # first two subsets
cvrt <- (findpoolreg::GMST %>% dplyr::filter(Year >= 1947, Year <= 2021))$smoothedGMST
```
Further, we need to assemble the centre points of the grid tiles in a matrix. As we can see, the centers are already contained in 
the grid data in the correct order, so we only transform the relevant columns to a matrix
```{r}
head(example_grid)
grid_centers <- as.matrix(example_grid[ , c("meanlon", "meanlat")])
```


Now we are ready to the function $\texttt{bootstrap_scalegev_subsets}$, which is the one 
that performs the max-stable-process-based bootstrap procedure.

```{r}
set.seed(1)
bootres <- bootstrap_scalegev_subsets(data = example_data, temp.cov = cvrt,
                                      locations = grid_centers, subsets = subsets, B = 500)
bootres
```
 
Once the bootstrapped p-values are computed, we can compute adjusted p-values with the function $\texttt{get_adj_pvals}$.
Here, we choose the Benjamini Hochberg and the Benjamini Yekutieli methods for adjusting the p-values. 

```{r}
bootres <- get_adj_pvals(bootres, methods = c("BH", "BY"))
bootres %>% dplyr::arrange(p_boot)
```

We can also plot the grid tiles shaded in the colours of the respective p-values, with the function $\texttt{visualise_test_res}$.
The method for adjusting p-values is provided in the argument method.
```{r}
visualise_test_res(coord_grid = example_grid, testres = bootres, method = "BH",
                   plot_type = "pvals", level = 0.1, Zoom = 2, position = "topright",
                   width = 800, height = 400)
```

Perfect, the test detects all deviations from the null-hypothesis.
We can also perform the boostrap based on bivariate extreme value distributions. 
For this, we apply the function $\texttt{bootstrap_pairs_bivmod}$ to the pairs defined above, or we 
just supply the location of primary interest, which is 5 in our case, and then all pairs will be tested
automatically.

```{r}
set.seed(1)
boot_biv <- bootstrap_pairs_bivmod(data = example_data, temp.cov = cvrt, loi = 5, B = 300)
boot_biv <- get_adj_pvals(boot_biv)
boot_biv
```

Again, the p-values can be visualised. This time, we want to visualise the p-values obtained when 
ignoring the multiple testing procedure, i.e. the plain bootstrap p-values.

```{r}
visualise_test_res(coord_grid = example_grid, testres = boot_biv, method = "boot",
                   plot_type = "pvals", level = 0.1, Zoom = 2, position = "topright",
                   width = 800, height = 400)
```
Again, the test detects all deviations.
Let's also try the more conservative Holm procedure. 

```{r}
visualise_test_res(coord_grid = example_grid, testres = boot_biv, method = "holm",
                   plot_type = "pvals", level = 0.1, Zoom = 2, position = "topright",
                   width = 800, height = 400)
```


