---
title: "Find pooling region"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Find pooling region}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(findpoolreg)
library(dplyr)
```

We are going to apply the bootstrap procedures on the example data set contained in this package. 
First, we load the data and grid coordinates.
```{r}
data("example_grid")
data("example_data")
```
We can have a look at the grid to get an impression of the region.
```{r}
leaflet::leaflet(data = example_grid)  %>% 
  leaflet::addTiles() %>%
  leaflet::addRectangles(lng1 = ~from_lon, lng2 = ~to_lon, lat1 = ~from_lat, lat2 = ~to_lat) %>% 
  leaflet::addLabelOnlyMarkers(~meanlon, ~meanlat, label =  ~as.character(Region), 
                               labelOptions = leaflet::labelOptions(noHide = TRUE, textOnly = TRUE,
                               textsize = 5, style = list("color" = "blue",  "font-weight" = "bold", "font-size" = "30px")))
```

Now, let's assume we want to find all regions for which the parameters of the scale-GEV distribution seem
sufficiently similar to those of the region with the label 5. 
To do so, we first apply the bootstrap procedure based on max-stable processes to all pairs \eqn{(5, d), d = 1, \ldots, 4, 6, \ldots, 9}.
We first make a list containing all those pairs and get values of the temporal covariate (here: 4-year smoothed GMST anomaly). 
For this, we assume that our data are observations from Years 1947 to 2021. 
Also, we check whether the column names of the data matrix are those used when defining the pairs (i.e. numbered from 1 to 9). 
```{r}
colnames(example_data)
subsets <- purrr::map(c(1:4, 6:9), ~ c(5, .x))
subsets[1:2]  # first two subsets
cvrt <- (findpoolreg::GMST %>% dplyr::filter(Year >= 1947, Year <= 2021))$smoothedGMST
```
Further, we need to assemble the centre points of the grid tiles in a matrix. As we can see, the centers are already contained in 
the grid data in the correct order, so we only transform the relevant columns to a matrix
```{r}
head(example_grid)
grid_centers <- as.matrix(example_grid[ , c("meanlon", "meanlat")])
```


Now we are ready to the function \code{bootstrap_scalegev_subsets}, which is the one 
that performs the max-stable-process-based bootstrap procedure.

```{r}
bootres <- bootstrap_scalegev_subsets(data = example_data, temp.cov = cvrt, locations = meancoords, subsets = subsets, B = 10)
bootres
```
 
Once the bootstrapped p-values are computed, we can compute adjusted p-values with the function \code{get_adj_pvals}.

```{r}
bootres <- get_adj_pvals(bootres, methods = c("BH", "BY"))
bootres %>% tidyr::unnest(cols = sbst) %>% dplyr::arrange(p_boot)
```



